---
title: "p8105_hw6_sjz2120"
author: "Sophie Zhang (sjz2120)"
date: "2022-11-22"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(patchwork)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
set.seed(1)
```



# Problem 1
For this problem, we’ll use the 2017 Central Park weather data that we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.

The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response and tmin as the predictor, and are interested in the distribution of two quantities estimated from these data:

* r̂ 2
* log(β̂ 0∗β̂ 1

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1). Note: broom::glance() is helpful for extracting r̂ 2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 0∗β̂ 1).



# Problem 2
The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository. Let's load in the raw dataset:

```{r Load-in-homicides-data, message=FALSE}
wp_homicides_url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

homicides =
  read_csv(wp_homicides_url,
           col_types = c("c", "c", "c", "c", "c", "n", "c", "c", "c", "d", "d", "c")) %>% # Read in the raw csv dataset
  janitor::clean_names() %>%
  mutate(victim_age = as.numeric(victim_age),
         reported_date = as.character(reported_date),
         reported_date = as.Date(reported_date, "%Y%m%d")) # Change variable types for `reported_date` and `age`

homicides
```

### Cleaning the `homicides` dataset
Now let's clean the `homicides` dataset. Below, I've created a `city_state` variable (combining the `city` and `state` variables), and a binary variable `solved` indicating whether the homicide is solved. Additionally, I've omitted the cities of Dallas, TX; Phoenix, AZ; and Kansas City, MO, because they don’t report victim race. I've also omitted Tulsa, AL since it's a data entry mistake.

For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r cleaning-homicides-dataset}
homicides_df =
  homicides %>%
  mutate(city_state = paste(city, state, sep = ", "),  # Create `city_state` and `solved` variables
         solved = ifelse(disposition %in% c("Closed by arrest", "Closed without arrest"), 1, 0)) %>%
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")), # Omit cities without data for race, and Tulsa AL
         victim_race %in% c("White", "Black")) # Limit analyses to those whose `victim_race` is white or Black
  
homicides_df
```

The cleaned and limited `homicides_df` dataset contains `r nrow(homicides_df)` rows/observations and `r ncol(homicides_df)` columns/variables. Each of the `r nrow(homicides_df)` rows represents a homicide. The `r ncol(homicides_df)` key variables in this dataset include: `r colnames(homicides_df)`:

* `uid` -- a unique ID for each observation
* `reported_date` -- the reported date of each homicide (yearmonthday)
* `victim_last` and `victim_first` -- the last and first name of each homicide victim
* Demographic information for each homicide victim, including their race (`victim_race`), age (`victim_age`), sex (`victim_sex`)
* The location of each homicide incident, including the `city` and `state` as well as the latitude and longitude (`lat` and `lon`)
* `disposition` -- the status of each homicide case (case closed/open and arrest status)
* `city_state` -- combines the `city` and `state` variables
* `solved` -- indicates whether the homicide is solved (1 if solved, 0 if unsolved)

### Solving Homicides in Baltimore, MD
For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r GLM-for-Baltimore}
# Creating a dataframe with only Baltimore, MD homicides data
baltimore_df =
  homicides_df %>%
  filter(city_state == "Baltimore, MD")

# Let's run a GLM on solved vs unsolved homicides in Baltimore, MD, calculate the OR and CI estimates, and save it as an R object `baltimore_glm`
baltimore_glm =
  baltimore_df %>%
  glm(solved ~ victim_age + victim_sex + victim_race,
    data = .,
    family = binomial()) %>%
  broom::tidy() %>%
  janitor::clean_names() %>%
  mutate(OR = exp(estimate),
         CI_lower = exp(estimate - 1.96*std_error),
         CI_upper = exp(estimate + 1.96*std_error))

# Not necessary, but I created a slightly neater table showing the OR and CI estimates for Baltimore
baltimore_tbl =
  baltimore_glm %>%
  select(term, OR, CI_lower, CI_upper) %>%
  knitr::kable(digits = 3)

baltimore_tbl
```


### Solving homicides in each city
Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r GLM-for-all-cities}
homicides_glm =
  homicides_df %>%
  nest(location = -city_state) %>%
  mutate(glm_cities = map(.x = location, ~glm(solved ~ victim_age + victim_sex + victim_race, data = homicides_df, family = binomial())),
         glm_tidy = map(glm_cities, broom::tidy),
         glm_results = map(glm_tidy, janitor::clean_names)) %>%
  select(city_state, glm_results) %>%
  unnest(glm_results) %>%
  mutate(OR = exp(estimate),
         CI_lower = exp(estimate - 1.96*std_error),
         CI_upper = exp(estimate + 1.96*std_error)) %>%
  select(city_state, term, OR, CI_lower, CI_upper)
  
homicides_glm
```


### Plot visualizing solving homicides for each city
Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r plotting-homicides}
homicides_glm %>%
  mutate(city_state = fct_reorder(city_state, OR)) %>%
  ggplot(aes(x = city_state, y = OR)) +
  geom_point()
```



# Problem 3
In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

