---
title: "p8105_hw6_sjz2120"
author: "Sophie Zhang (sjz2120)"
date: "2022-11-22"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(patchwork)
library(modelr)
library(mgcv)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
set.seed(1)
```



# Problem 1
For this problem, we’ll use the 2017 Central Park weather data that we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.

The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response and tmin as the predictor, and are interested in the distribution of two quantities estimated from these data:

* r̂ 2
* log(β̂ 0∗β̂ 1

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1). Note: broom::glance() is helpful for extracting r̂ 2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 0∗β̂ 1).



# Problem 2
The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository. Let's load in the raw dataset:

```{r Load-in-homicides-data, warning=FALSE}
wp_homicides_url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

homicides =
  read_csv(wp_homicides_url,
           col_types = c("c", "c", "c", "c", "c", "n", "c", "c", "c", "d", "d", "c")) %>% # Read in the raw csv dataset
  janitor::clean_names()

homicides
```

### Cleaning the `homicides` dataset
Now let's clean the `homicides` dataset. Below, I've created a `city_state` variable (combining the `city` and `state` variables), and a binary variable `solved` indicating whether the homicide is solved. Additionally, I've omitted the cities of Dallas, TX; Phoenix, AZ; and Kansas City, MO, because they don’t report victim race. I've also omitted Tulsa, AL since it's a data entry mistake.

For this problem, I've limited my analysis to those for whom victim_race is white or Black.

```{r cleaning-homicides-dataset, warning=FALSE}
homicides_df =
  homicides %>%
  mutate(city_state = paste(city, state, sep = ", "),  # Create `city_state` and `solved` variables
         solved = ifelse(disposition == "Closed by arrest", 1, 0)) %>%
  filter(!(city_state %in% c("Dallas, TX",
                             "Phoenix, AZ",
                             "Kansas City, MO",
                             "Tulsa, AL")), # Omit cities
         victim_race %in% c("White", "Black")) %>% # Limit analyses
  mutate(victim_age = as.numeric(victim_age)) %>% # Change variable types
  select(victim_race, victim_age, victim_sex, city_state, solved) # Keep only the relevant variables
  
homicides_df
```

The cleaned and limited `homicides_df` dataset contains `r nrow(homicides_df)` rows/observations and `r ncol(homicides_df)` columns/variables. Each of the `r nrow(homicides_df)` rows represents a homicide. The `r ncol(homicides_df)` key variables in this dataset include: `r colnames(homicides_df)`:

* Demographic information for each homicide victim, including their race (`victim_race`), age (`victim_age`), sex (`victim_sex`)
* `city_state` -- combines the `city` and `state` variables
* `solved` -- indicates whether the homicide is solved (1 if solved, 0 if unsolved)


### Solving Homicides in Baltimore, MD
For the city of Baltimore, MD, I used the GLM function to fit a logistic regression with resolved vs unresolved homicides (as the outcome) and victim age, sex, and race as predictors. The GLM output has been saved as `baltimore_glm`, and has been tidied. Additionally, I've created columns in `baltimore_glm` for the estimate and confidence interval of the adjusted odds ratio for solving homicides, keeping all other variables fixed.

```{r GLM-for-Baltimore}
# Creating a dataframe with only Baltimore, MD homicides data
baltimore_df =
  homicides_df %>%
  filter(city_state == "Baltimore, MD")

# Let's run a GLM on solved vs unsolved homicides in Baltimore, MD, calculate the OR and CI estimates, and save it as an R object `baltimore_glm`
baltimore_glm =
  baltimore_df %>%
  glm(solved ~ victim_age + victim_sex + victim_race,
    data = .,
    family = binomial()) %>%
  broom::tidy() %>%
  janitor::clean_names() %>%
  mutate(OR = exp(estimate),
         CI_lower = exp(estimate - 1.96*std_error),
         CI_upper = exp(estimate + 1.96*std_error))

baltimore_glm
```

Now, let's obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims, keeping all other variables fixed.
```{r homicides-baltimore-gender}
baltimore_glm %>%
  filter(term == "victim_sexMale") %>%
  select(OR, CI_lower, CI_upper) %>%
  knitr::kable(col.names = c('Odds Ratio', 'CI (lower)', 'CI (upper)'))
```
* The adjusted odds ratio for solving homicides comparing male victims to female victims (keeping all other variables fixed) is OR = `r baltimore_glm %>% filter(term == "victim_sexMale") %>% pull(OR)` (95% CI: `r baltimore_glm %>% filter(term == "victim_sexMale") %>% pull(CI_lower)`, `r baltimore_glm %>% filter(term == "victim_sexMale") %>% pull(CI_upper)`).


### Solving homicides in each city
Below, I've run a GLM for each of the cities in the `homicides_df` dataset and saved the output as `homicides_glm_df`. I've additionally extracted the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims (keeping all other factors constant) for each city, and have saved this more limited dataframe as `homicides_glm_gender_df`.

```{r GLM-for-all-cities}
homicides_glm_df =
  homicides_df %>%
  nest(data = -city_state) %>%
  mutate(glm_cities = map(.x = data, ~glm(solved ~ victim_age + victim_sex + victim_race, data = .x, family = binomial())),
         glm_tidy = map(glm_cities, broom::tidy)) %>%
  select(city_state, glm_tidy) %>%
  unnest(glm_tidy) %>%
  mutate(OR = exp(estimate),
         CI_lower = exp(estimate - 1.96*std.error),
         CI_upper = exp(estimate + 1.96*std.error)) %>%
  select(city_state, term, OR, CI_lower, CI_upper)

homicides_glm_gender_df =
  homicides_glm_df %>%
  filter(term == "victim_sexMale") %>%
  select(city_state, OR, CI_lower, CI_upper)
```


### Plot visualizing solving homicides for each city
Below, I've created a plot called `homicides_gender_plot` that shows the estimated ORs and CIs for solving homicides comparing male victims to female victims in each city. The cities are organized according to estimated OR.

```{r plotting-homicides}
homicides_gender_plot =
  homicides_glm_gender_df %>%
  mutate(city_state = fct_reorder(city_state, OR)) %>%
  ggplot(aes(x = city_state, y = OR)) +
  geom_point(alpha = 0.5) +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper),
                alpha = 0.5) +
  labs(title = "Estimated odds ratios & 95% confidence intervals of solved homicides, by city",
        subtitle = "ORs comparing male vs female homicide victims",
        x = "City (City, State)",
        y = "Estimated ORs for solved homicides, male vs female (w/95% CI)") +
  theme(plot.title = element_text(size = 11, hjust = 0.5),
        plot.subtitle = element_text(size = 10, hjust = 0.5),
        axis.text.x = element_text(size = 7, angle = 60, hjust = 1),
        axis.text.y = element_text(size = 7, hjust = 0.5),
        legend.position = "none")

homicides_gender_plot
```

#### Describing the solved homicides comparing victim sex

From the `homicides_gender_plot` above, we see that New York, NY has the lowest adjusted OR for solving homicides comparing male victims to female victims keeping all other variables fixed; while Albuquerque, NM has the highest.

An OR = 1 would mean that there is no association between victim sex and homicide solve status (i.e. no disparity in homicide case solving among the sexes). However, we see that most of the cities have an adjusted OR < 1 for solving homicides comparing male victims to female victims, indicating that homicide cases involving male victims have lower odds of being solved compared to cases involving female victims. On the other hand, cities with ORs higher than 1 (including Nashville, TN; Fresno, CA; Stockton, CA; and Albuquerque, NM) are cities where homicide cases involving male victims have higher odds of being solved compared to cases involving female victims.


# Problem 3
Now let's analyze data gathered to understand the effects of several variables on a child’s birthweight. The dataset `birthweight.csv` is located in the `data` folder in this RProject. It consists of roughly 4000 children and includes many variables related to maternal and newborn status measured during childbirth.

### Loading and cleaning the `birthweight_df` dataset
Below, I've loaded in and cleaned the data for regression analysis, converting numeric to factor where appropriate. The dataset is saved as `birthweight_df`

```{r Load-and-clean-birthweight-data, message=FALSE}
# Let's load in the dataset!
birthweight_df =
  read_csv("data/birthweight.csv") %>%
  janitor::clean_names() %>%
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         malform = as.factor(malform),
         mrace = as.factor(mrace))

birthweight_df
```

Now let's finish cleaning the dataset and check for missing data in `birthweight_df`! There seems to be 0 missing data for all of the variables in the dataframe:)

```{r birthweight-check-missing}
# Let's check for missing data:
map(birthweight_df, ~sum(is.na(.)))
```


### Describing the `birthweight_df` dataset
The cleaned `birthweight_df` dataset contains `r nrow(birthweight_df)` rows/observations and `r ncol(birthweight_df)` columns/variables. After a quick check, we also see that the dataset has no missing data (see above). Each of the `r nrow(birthweight_df)` rows represents a birth The `r ncol(birthweight_df)` key variables in this dataset include: `r colnames(birthweight_df)`:

* `babysex`: baby’s sex (male = 1, female = 2)
* `bhead`: baby’s head circumference at birth (centimeters)
* `blength`: baby’s length at birth (centimeteres)
* `bwt`: baby’s birth weight (grams)
* `delwt`: mother’s weight at delivery (pounds)
* `fincome`: family monthly income (in hundreds, rounded)
* `frace`: father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
* `gaweeks`: gestational age in weeks
* `malform`: presence of malformations that could affect weight (0 = absent, 1 = present)
* `menarche`: mother’s age at menarche (years)
* `mheigth`: mother’s height (inches)
* `momage`: mother’s age at delivery (years)
* `mrace`: mother’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
* `parity`: number of live births prior to this pregnancy
* `pnumlbw`: previous number of low birth weight babies
* `pnumsga`: number of prior small for gestational age babies
* `ppbmi`: mother’s pre-pregnancy BMI
* `ppwt`: mother’s pre-pregnancy weight (pounds)
* `smoken`: average number of cigarettes smoked per day during pregnancy
* `wtgain`: mother’s weight gain during pregnancy (pounds)


### Let's make a regression model for birthweight!
Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two.

A quick Google search of factors which affect birthweight (including the Wikipedia page [https://en.wikipedia.org/wiki/Birth_weight#Determinants] and the articles cited by the page) suggest that factors such as maternal age at birth (`momage`), number of prior births (`parity`), and smoking during pregnancy (`smoken`) have all been found to be significantly associated with birthweight. Additionally, sociodemographic factors such as race (`mrace`) -- and specifically increased stress caused by racism -- have also been found to be associated with birthweight. 

That's why for `model_1` below I chose to include the variables `momage`, `parity`, `smoken`, and `mrace` as predictors of birthweight. I chose not to include an interaction term between any of the variables, since I wasn't able to find much data suggesting an interaction between any of the chosen predictors in relation to birthweight.

```{r model-1}
# Model 1: Monthly income, gestational age, mother's age at delivery, number of live births prior to this pregnancy, number of prior small for gestational age babies, mother's pre-pregnancy BMI
model_1 = lm(bwt ~ momage + parity + smoken + mrace, data = birthweight_df)
model_1
```

### Residual plot for our (first) birthweight model
Below, I've created a plot of model residuals against fitted values for the birthweight regression model previously created. I used `add_predictions` and `add_residuals` from the `modelr` package to make this plot.

```{r model-1-residual-plot}
residual_plot =
  birthweight_df %>%
  add_predictions(model_1) %>%
  add_residuals(model_1) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.3, size = 0.3) +
  geom_smooth(se = FALSE, method = "lm") +
  labs(title = "Model 1 Residuals Plot",
      x = "Predicted (Fitted Values)",
      y = "Residuals") +
  theme(plot.title = element_text(size = 11, hjust = 0.5),
        plot.subtitle = element_text(size = 10, hjust = 0.5),
        axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 7, hjust = 0.5),
        legend.position = "none")

residual_plot
```

There seem to be two more dense clusters, but otherwise the residuals seem to be centered around 0 and spread evenly on either side of the zero line.


### Creating more regression models for birthweight
Let's fit two other linear regression models:

* **Model 2**: Using length at birth and gestational age as predictors (main effects only)
* **Model 3**: Using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r other-models}
# Model 2: Length at birth and gestational age as predictors (main effects only)
model_2 = lm(bwt ~ blength + gaweeks, data = birthweight_df)

# Model 3: Head circumference, length, sex, and all interactions (including the three-way interaction)
model_3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = birthweight_df)

model_2
model_3
```


### Cross validation of our regression models
Let's compare model 1 to these models 2 and 3! We'll make this comparison in terms of the cross-validated prediction error, using `crossv_mc` and functions in `purrr` (see code chunk below).

```{r cross-validating-models, warning=FALSE}
# Initial setup for cross validation
cv_df =
  crossv_mc(birthweight_df, 100) %>%
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))

# Cross validating our 3 models
cv_df = 
  cv_df %>%
  mutate(
    model_1 = map(train, ~lm(bwt ~ momage + parity + smoken + mrace, data = .x)),
    model_2 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    model_3 = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = .x))
    ) %>%
  mutate(
    rmse_model_1 = map2_dbl(model_1, test, ~rmse(model = .x, data = .y)),
    rmse_model_2 = map2_dbl(model_2, test, ~rmse(model = .x, data = .y)),
    rmse_model_3 = map2_dbl(model_3, test, ~rmse(model = .x, data = .y))
    )
```

Now we can compare the 3 models using a violin plot:
```{r compare-models-plot}
compare_models_plot =
  cv_df %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin() +
  labs(title = "Comparing Our Models for Birthweight",
      x = "Model",
      y = "rmse") +
  theme(plot.title = element_text(size = 12, hjust = 0.5))

compare_models_plot
```

*Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.*

Thank goodness for this because `model_1` (the model I created based on a purely hypothesized structure for the factors that underly birthweight) has the largest rmse among all the models, suggesting that it's the worst performing out of all of the linear models in terms of predictive accuracy 😞
